CASTRO: A New Compressible Astrophysical Solver. IV. Performance Portability and GPUs

### 2019-03-22 gpus/paper.tex
more writing

\newcommand{\castro}{{\sf Castro}}
\newcommand{\maestro}{{\sf Maestro}}

@@

\begin{document}
%======================================================================
% Title
%======================================================================
\title{CASTRO: A New Compressible Astrophysical Solver. IV. Performance Portability}

\shorttitle{}
\shortauthors{}

@@

GPU-enabled integration has been explored by \citep{brock:2015}.

In this paper, we describe our approach to performance portability
with the \castro\ multiphysics simulation code..  Our goal is to have
the same computational kernels run efficiently on CPUs (including
manycore processors) and GPUs.  We describe the work we needed to do
to enable this portability and show performance metrics for a reacting
flow problem in astrophysics.

@@

\castro\ is a compressible (radiation-)hydrodynamics simulation code built on
the AMReX adaptive mesh refinement framework \citep{castro,castro2,castro3}.  It 
includes full self-gravity using multigrid methods with isolated boundaries
\citep{katz:2016}, works with general equations of state \citep{zingalekatz}
and arbitrary nuclear reaction networks.  Castro has been used for science investigations of
...

The original computing architectures targeted were

\castro\ development began with \cite{castro}, with the goal of
creating an astrophysical hydrodynamics code using a modern adaptive
mesh refinement library (BoxLib at the time) and unsplit hydrodynamics
techniques.  The initial application area was nuclear astrophysics,
including core-collapse supernovae, so it was designed to work with a
general equation of state~\cite{zingalekatz} and reaction network, and
with radiation transport (through flux limited diffusion,
\citealt{castro2,castro3}).  \castro\ also acted as the compressible
counterpart to our low Mach number hydrodynamics code \maestro,
sharing much of the same infrastructure, and allowing for simulations
to be restarted from the low Mach regime into the compressible
regime~\cite{malone:2014}.  \castro\ has been applied to various
progenitor models of Type Ia
supernova~\cite{malone:2014,katz:2016,polin:2019}, exoplanet
dynamics~\cite{ryu:2018}, core-collapse and population III
supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray bursts~\cite{astronum:2018}..

@@

@@

@@

\begin{figure*}
\centering
\includegraphics[width=0.3\textwidth]{gpu_1} 
\includegraphics[width=0.3\textwidth]{gpu_2} 
\includegraphics[width=0.3\textwidth]{gpu_3}
\caption{\label{fig:loops} caption}
\end{figure*}

@@

Future work: gravity on GPUs (multigrid or multipole), port of \maestro\ following
the same ideas.

@@

\acknowledgements \castro\ is freely available at
\url{http://github.com/AMReX-Astro/Castro}.  The work at Stony Brook
was supported by DOE/Office of Nuclear Physics grant DE-FG02-87ER40317
and NSF award AST-1211563.  An award of computer time was provided by
the Innovative and Novel Computational Impact on Theory and Experiment
(INCITE) program.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725.  We thank NVIDIA Corporation
for the donation of a Titan X Pascal and Titan V used in this research.

>>>>

\newcommand{\castro}{{\sf Castro}}
\newcommand{\maestro}{{\sf Maestro}}
\newcommand{\maestroex}{{\sf MAESTROeX}}

@@

\newcommand{\fab}{{\sf Fab}}
\newcommand{\multifab}{{\sf MultiFab}}

\begin{document}
%======================================================================
% Title
%======================================================================
\title{CASTRO: A New Compressible Astrophysical Solver. IV. Performance Portability and GPUs}

\shorttitle{CASTRO Performance Portability and GPUs}
\shortauthors{}

@@

GPU-enabled integration has been explored by \citep{brock:2015}.\MarginPar{XNet paper?}

In this paper, we describe our approach to performance portability
with the \castro\ multiphysics simulation code.  Our goal is to have
the same computational kernels run efficiently on CPUs (including
manycore processors) and GPUs.  We describe the work we needed to do
to enable this portability and show performance metrics for a reacting
flow problem in astrophysics.

@@

\castro\ development began with \cite{castro}, with the goal of
creating an astrophysical hydrodynamics code using a modern adaptive
mesh refinement library with space- and time-refinement (BoxLib at the
time) and unsplit hydrodynamics techniques.  \castro\ now uses the
AMReX adaptive mesh refinement library to handle the computational
grid, memory management, and parallelism, and the \castro\ and AMReX
developers work closely together to The initial application area was
nuclear astrophysics, including core-collapse supernovae, so it was
designed to work with a general equation of state~\cite{zingalekatz}
and arbitrary nuclear reaction networks, and with radiation transport
(through flux limited diffusion, \citealt{castro2,castro3}).  It
includes full self-gravity using multigrid methods with isolated
boundaries \citep{katz:2016}.  Finally, \castro\ also acted as the
compressible counterpart to our low Mach number hydrodynamics code
\maestro, sharing much of the same infrastructure, and allowing for
simulations to be restarted from the low Mach regime into the
compressible regime~\cite{malone:2014}.

\castro\ has been used for science investigations of various different
progenitor models of Type Ia
supernova~\cite{malone:2014,katz:2016,polin:2019}, exoplanet
dynamics~\cite{ryu:2018}, core-collapse and population III
supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray
bursts~\cite{astronum:2018}..

@@

In \castro\ the domain is decomposed into a number of grids at various
levels of refinement.  A single grid and its associated data is called
a \fab, and its location in space is specified uniquely by the index
of its lower left cell, typically called {\tt lo}, and its upper right
cell, {\tt hi}, in a global index space for that level of refinement.
The collection of all the \fab s at a given level of refinement is
called a \multifab.  Time advancement proceeds by advancing all of the
grids at the coarsest level through a timestep $\Delta t$, and then
the grids at the next refinement level through {\tt rr} timesteps of
$\Delta t/\mathtt{rr}$, where {\tt rr} is the refinement ratio between
the levels, typically 2 or 4.  At that point, the two levels are at
the same time and a synchronization is done, as discussed in \cite{castro}.

\begin{figure*}[t]
\centering
\includegraphics[width=0.33\textwidth]{gpu_1} 
\includegraphics[width=0.33\textwidth]{gpu_2} 
\includegraphics[width=0.33\textwidth]{gpu_3}
\caption{\label{fig:loops} caption}
\end{figure*}

The main drivers in Castro, including the memory management and
parallelism are written in C++.  The vast majority of the compute
kernels are written in modern Fortran, although with time we are
converting these to C++ as well.

@@

GPUs achieve their performance by having a large number of threads
(1000s) that can execute the same instructions on multiple pieces of
data (a SIMD approach). \MarginPar{Max will not like this desciption.}
In modern supercomputers, like Summit at OLCF, more than 95\% of the
theoretical floating point performance is from the
GPUs \MarginPar{cite?}.  The main bottleneck with these hybrid systems
is copying the data back and forth between the main memory accessible
to the CPU and the memory on the GPU.  Therefore, to get the best
perfomance, we want to copy the data once onto the GPUs, operate as
much as possible on it while it is on the GPUs, and then copy back to
CPUs only when needed.  To simplify this data transfer, we use CUDA
managed memory.\MarginPar{more here}.  This approach will require us
to write all of our code in a manner that can be run on the GPUs.  We
discuss some of the issues here in this section.

There are several approaches to offloading a compute kernel to GPUs,
including the directive-based OpenACC and OpenMP, and the
NVIDIA-specific CUDA library.  For the present work, we use CUDA,
although our approach can mix the use of the directive-based
offloading instead of using pure CUDA.  This is something we will
explore in the future.

@@

@@

Self-gravity is the next important physics module that needs to be
developed for GPUs.  \castro\ uses either geometric multigrid,
implemented via AMReX, or a multipole expansion to solve for the
gravitational potential.  With Strang splitting of the reactions, it
is possible to asynchronously solve for the gravitational potential on
CPUs while solving the second-half of the reaction system on GPUs,
since the operator split approach does not change density via
reactions.  This is the approach we will explore in the near future
for self-gravitating reacting flows.

We will also begin using the same ideas described here to port the
hydrodynamics and reactions in \maestroex\ to GPUs.  The projection of
the velocity field will remain on CPUs in the near future.

@@

\acknowledgements \castro\ is available at
\url{http://github.com/AMReX-Astro/Castro}.  The work at Stony Brook
was supported by DOE/Office of Nuclear Physics grant DE-FG02-87ER40317
and NSF award AST-1211563.  An award of computer time was provided by
the Innovative and Novel Computational Impact on Theory and Experiment
(INCITE) program.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725.  In particular, we thank
OLCF for early access to the Summitdev platform, where we could
develop and debug our GPU approach.  We also thank NVIDIA Corporation
for the donation of a Titan X Pascal and Titan V used in this
research.  The GPU development of \castro\ benefited greatly from
numerous GPU hackathons arranged by OLCF.
